### 2022-10-09 dissertation/chapters/01-introduction.tex
✨ (dissertation): Chapter 1 done

\begin{quotation}
      \textit{``It is better to solve one problem five different ways, than to solve five problems one way.'' - George Polya}
\end{quotation}

\Ac{ML} is one of the most popular fields of research in \ac{AI} studies today. In recent years, \ac{ML} research has seen some notable achievements in academia \cite{ref:lecun:2015, ref:glorot:2010, ref:goodfellow:2014, ref:quoc:2017}, as well as the industry at large \cite{ref:silver:2016, ref:silver:2017, ref:zoph:2017, ref:lewis:2017}.  \ac{ML} research has grown tremendously over the past decade with successes like AlphaGo, which set new standards for \ac{AI} capabilities by beating the world's best Go player, Lee Sedol, 4-1 \cite{ref:san-hun:2016}.  Furthermore, an improvement on AlphaGo, called AlphaZero, learned to play Go  \textit{tabula rasa} \footnote{Latin word meaning to learn from no prior knowledge.} and managed to beat AlphaGo, 100-1 \cite{ref:silver:2017}.

Over the past few years, modern hardware capabilities have improved to the point where workloads in the field of \ac{ML} that were previously computationally infeasible, are now possible. One such sub-field of \ac{ML} is \acp{ANN}. \acp{ANN} can generally be described as well-organised structures of mathematical computation and are inspired by the biological brain \cite{ref:engelbrecht:2007}. \acp{ANN} can be trained, which is the equivalent of ``learning'' from data. Learning gives rise to the ability to apply some form of decision making. This ability provides a wide-range of applications from healthcare to finance and yields great interest in the field. With the improvement of hardware came an influx of \ac{ML} researchers that focused their attention on the training of \acp{ANN}.

\Acfp{FFNN} are specific types of \acp{ANN} \cite{ref:reed:1999}. A popular field of focus for studying \acp{ANN} is the process by which these models are trained.  The most common way of training \acp{FFNN} is \index{supervised learning}\textit{supervised learning}, which is a training technique that involves exposing the \acp{ANN} with input and comparing the produced output to that of predefined target data. Training of \acp{ANN} is seen as an optimisation problem. The \ac{ANN} maintains a set of parameters, referred to as ``weights'' and ``biases''. A search algorithm known as a \index{heuristic}\textit{heuristic} \cite{ref:pearl:1984} is used to assign optimal values to the parameters of the \ac{ANN}, such that a specified objective function is minimised.

This dissertation focuses on the development of a specific type of heuristic, called a \acl{HH} to train \acp{FFNN} in a supervised learning approach. In particular, this chapter provides the reader with a brief overview of the problem domain and outlines the research problem, objectives and purpose. The remainder of the chapter is presented as follows:

\begin{itemize}
      \item \textbf{Section \ref{sec:introduction:summary_research_domain}} provides the reader with brief summary of the problem domain. This section also provides a review of the history that lead to related \ac{ANN} research.

      \item \textbf{Section \ref{sec:introduction:problem}} outlines the research problem being addressed in this dissertation.

      \item \textbf{Section \ref{sec:introduction:motivation}} provides motivations as to why this research is important and valuable to the field.

      \item \textbf{Section \ref{sec:introduction:objectives}} presents the reader with research objectives and outlines the goals of this dissertation.

@@

      \item Finally, a full dissertation outline is provided in \textbf{Section \ref{sec:introduction:outline}}.
\end{itemize}

The reader is presented with the summary of the research domain next.

@@

In order to grasp all the concepts that are included in this disseration, it is important to briefly review the landscape of the research fields that form the basis for the research in this disseration. This section provides the reader with a brief summary of the research field and outlines key concepts that contribute to the topics addressed and discussed in this disseration.

Historically, \ac{ML} models were built with a specific purpose in mind and revolved around specialised applications. A notable development was that of Deep Blue by IBM \cite{ref:keene:1996}, which managed to play chess at a grand master level. Deep Blue was able to finish 2-4 against the greatest chess player at the time, Gary Kasparov. Although Deep Blue lost the tournament, this was a major breakthrough in the field of \ac{AI}. However, Deep Blue was only able to play chess and could not be utilised for any other purposes \cite{ref:kelley:2010}. Today, DeepMind's \ac{RL} algorithm is capable of playing all of the original Atari games \cite{ref:mnih:2013}, but is still very limited to a small set of problems that it can solve. Although the computational requirements were lagging behind, these achievements where impressive at the time.

@@

Generalisation of \acp{ANN} refer to the capabilities of the network to either perform well on unseen data or to be applied to other problem domains. Various attempts have been made to improve the generalisation capabilities of \acp{ANN}. Examples of techniques used include weight \index{dropout}dropout \cite{ref:srivastava:2014}, \index{weight decay}weight decay \cite{ref:krogh:1992}, \index{meta-learning}meta-learning of learning parameters such as the \index{learning rate}learning rate and \index{momentum}momentum of gradient-based heuristics \cite{ref:zeiler:2012, ref:lv:2017, ref:darken:1992}, and removing or postponing the use of outliers in the training data \cite{ref:reeves:1998}. Other techniques focus on improving generalisation through model design. Design configurations focus on the layout of the neurons and structure of the \acs{ANN} (architecture) and how neurons are connected to interact (topology). Examples of such techniques include learning of \ac{ANN} architectures through pruning techniques \cite{ref:cibas:1996, ref:engelbrecht:1996} or constructive techniques \cite{ref:hassibi:1994, ref:lecun:1990}, adaptive activation functions \cite{ref:engelbrecht:1995, ref:fletcher:1994}, and quantum methods \cite{ref:wan:2017, ref:ricks:2004}.

@@

An example of dynamically balancing exploitation and exploration during the search process is by dynamically adjusting and learning the heuristic \index{hyper-parameters}\textit{hyper-parameters} as part of the learning process. This field of study is known as \index{meta-learning}meta-learning \cite{ref:giraud:2004}. \index{meta-learning}Meta-learning of heuristic hyper-parameters as applied in the training of \acp{ANN} has shown to yield good generalisation results \cite{ref:hospedales:2020, ref:vilalta:2002}.

The dynamic nature of the learning process leads towards the idea that \ac{ML} models could yield better generalisation capabilities if the learning process and learning mechanism applied are not statically defined, but rather dynamic and under the control of some mechanism. Furthermore, a learning technique that has at its disposal a more diverse set of learning behaviours from which it can dynamically select (when the need arises) could yield better generalisation capabilities \cite{ref:huang:2009}.

A recent suggestion related to the field of \index{meta-learning}meta-learning is to dynamically select and/or adjust the heuristic used throughout the training process. This approach focuses on the hybridisation of learning paradigms. The main concept behind this paradigm is that the learning process is dynamic and differs from problem to problem. A particular learning technique might work well for one problem and not for another. At the same time, a particular learning technique might work well for a particular part of the search landscape, but not for another. By dynamically combining the best of different learning paradigms throughout the learning process, a trade-off can be made between exploration and exploitation as is required.

One such form of hybridisation of learning paradigms is referred to as \index{heterogeneous learning approaches}\textit{heterogeneous learning approaches}. Heterogeneous learning approaches make use of different search behaviours by selecting from a behaviour pool. Heterogeneous approaches have shown to balance the trade-off between exploration and exploitation \cite{ref:nepomuceno:2013}.

A step further in the concept of hybridisation of learning paradigms is that of hybridisation of different heuristics as they are applied to some optimisation problem \cite{ref:burke:2013}. These methods are referred to as \acfp{HH} and focus on finding the best heuristic in \textit{heuristic-space} to solve a specific problem. One such form of \ac{HH} is a population-based approach that guides the search process by automatically selecting heuristics from a heuristic-space to be applied to a collection of different candidate solutions in the solution-space. This collection of candidate solutions are referred to as a \textit{population} of \textit{entities}, where each \textit{entity} is a single candidate solution to the problem being optimised.

Population-based \acp{HH} implement the strategy of multiple heuristics working together to solve a problem. However, finding the best heuristic to use is non-trivial and some \textit{selection strategy} must be used to select the best heuristic. \acp{HH} that implement such a selection strategy are referred to as \textit{selection} \acp{HH}. The term \textit{selection} refers to the ability of the \ac{HH} to select the best heuristic from a pool of low-level heuristics.

These selection \acp{HH} can be further categorised based on what information they use in the selection strategy. One specific type of selection \ac{HH} is called a \index{multi-method populated-based meta-heuristic}\textit{multi-method population-based \index{meta-heuristic}meta-heuristic} \cite{ref:vanderstockt:2018}. The term \index{multi-method}\textit{multi-method} refers to the incorporation of different low-level heuristics, with different search behaviours, into the heuristic-space. The term \textit{population-based} refers to the utilisation of a population of entities that represent candidate solutions. Finally, the term \index{meta-heuristic}\textit{meta-heuristics} refers to the \ac{HH} as a heuristic that does not have any domain knowledge and only makes use of information from the search process. \citeauthor{ref:grobler:2015} \cite{ref:grobler:2015} mentioned that \acp{HH} have been shown to solve a number of problems including bin-packing, examination timetabling, production scheduling, the traveling salesman problem, vehicle routing problem and many more.

The following section outlines the research problem that this dissertation aims to focus on and is followed by a discussion as to why this research endevour is important and valuable to the field.

@@

Many different heuristics have been developed and used to train \acp{FFNN} \cite{ref:gudise:2003, ref:rakitianskaia:2012, ref:montana:1989}. Each of these heuristics have different search behaviours, characteristics, strengths and weaknesses. Finding the best heuristic to train the \ac{FFNN} is required in order to yield optimal results. This process is often non-trivial and could be a time-consuming exercise.  Consider that selection of the best heuristic as applied to optimisation problems, such as training \acp{FFNN}, is very problem dependent \cite{ref:allen:1996, ref:drake:2020, ref:pillay:2018}.

Careful, systematic selection is thus required to find and select the best heuristic to train \acp{FFNN}. In the past, researchers selected the best heuristic by trail and error. A set of heuristics and carefully selected hyper-parameters would be implemented, followed by an empirical test to evaluate the performance of each heuristic for a given problem domain \cite{ref:pillay:2015}. In this way, researchers were able to determine which heuristics and which hyper-parameters performed well for different problems. However, this approach is problematic, because it is time-consuming and tedious.

The following section provides the authors' motivation for this research and outlines the value that this research brings to the field.

@@

Section \ref{sec:introduction:problem} provides the reader with the research problem being addressed in this disseration. This section provides compelling arguments for why this research is important and valuable to the field.

@@

In Section \ref{sec:introduction:problem} above, it is mentioned that finding the best heuristic to train \acp{FFNN} is a timely and tedious process. A modern approach is to use \acp{HH} to automate the process of selecting the best heuristic when applied to some optimisation problem. The best heuristic might not be a single heuristic, but rather a hybridisation of heuristics \cite{ref:pillay:2015}. Therefore, careful consideration is required to include and select from a diverse set of heuristics that is capable of solving the given problem. Importantly, only heuristics that are known to have the potential to solve the problem are applicable and can be included in the set of heuristics to select from.

In the general context of optimisation, many different types of \acp{HH} have been implemented and applied to many different problems. Some notable examples include the \index{simulated annealing}simulated annealing-based \ac{HH} by \citeauthor{ref:dowsland:2007} \cite{ref:dowsland:2007}, the \index{tabu-search}tabu-search \ac{HH} of \citeauthor{ref:burke:2010} \cite{ref:burke:2010}, the \index{heterogeneous meta-hyper-heuristic}heterogeneous meta-hyper-heuristic by \citeauthor{ref:grobler:2012} \cite{ref:grobler:2012} and work done by \citeauthor{ref:vanderstockt:2018} \cite{ref:vanderstockt:2018} on the analysis of selection hyper-heuristics for population-based metaheuristics in real-valued dynamic optimization. However, research on the application of \acp{HH} in the context of \acs{FFNN} training is scarce. \citeauthor{ref:nel:2021} \cite{ref:nel:2021} provides the first research in this field, applying a \acs{HH} to \acs{FFNN} training. The shortage of research in this field certainly warrants more exploration.

\acp{FFNN} training is seen as a computational search problem. \Acl{HH} as defined by Burke et al. \cite{ref:burke:2010} is ``a search method or learning mechanism for selecting or generating heuristics to solve computational search problems''. \acp{HH} is a field of research aimed at the automated selection, combination, adaptation or generation of multiple lower-level heuristics to efficiently solve computational search problems. This implies that it should be possible to apply \acp{HH} in the context \acs{FFNN} training.

@@

This research takes a particular interest in developing a high-level heuristic that makes use of probability theory and Bayesian statistical concepts to guide the heuristic selection process. This research develops the novel \Acf{BHH}, a new high-level heuristic that utilises a statistical approach, referred to as \index{Bayesian analysis}\textit{Bayesian analysis}, which combines prior information with new evidence to the parameters of a selection probability distribution. This selection probability distribution is the mechanism by which the \ac{HH} selects appropriate heuristics to train \acp{FFNN} during the training process (online learning).

@@

The main objective of this research is focused on developing a novel \Ac{BHH} selection mechanism in a \ac{HH} framework that can be used to train \acp{FFNN}. In order to reach this objective, the following sub-objectives are defined:

@@

      \item Conduct a literature study on different types of heuristics that have been used to train \acp{FFNN}. The literature study will provide the necessary background information to understand how these different heuristics can be used in the set of heuristics to be selected to train the \ac{FFNN}.

@@

      \item Conduct an empirical study to investigate the behavioural characteristics of the \ac{BHH} as it is used to train \acp{FFNN} over a number of different datasets.

      \item Conduct an empirical study and critically evaluate the performance of the developed \Ac{BHH} compared to individual heuristics in the heuristic-space as they are used to train \acp{FFNN} on different problems.

@@

The following section outlines the research contributions that this disseration yields.

@@

The results obtained from this research contribute to the field of study in the
following ways:

@@

      \item The results of the empirical study show that the \Ac{BHH} is able to select the best heuristic to train \ac{FFNN} in general, with statistical certainty. This relieves researchers from the burden of having to do this selection process manually through trial and error.

@@

The remainder of this dissertation is structured as follows:

\begin{itemize}
      \item \textbf{Chapter~\ref{chap:anns}} provides a literature study on \acp{ANN} and the various components that make up an \acl{AN}. The focus is on the training of \acp{FFNN}. It is shown how the training of \acp{FFNN} is seen as an optimisation problem.

@@

      \item \textbf{Chapter~\ref{chap:methodology}} presents a detailed description of the empirical process and the setup of each experiment. It discusses the datasets used, the \ac{FFNN} architecture and topology used, heuristics that are used, the configuration of hyper-parameters, initialisation techniques used, performance measures used and other smaller details of the implementation of each experiment.  Finally, discussions follow on how the results are analysed. This includes a discussion on the method used to determine statistical significance.

      \item \textbf{Chapter~\ref{chap:results}} provides and discusses the results of the empirical study, in detail. A baseline comparison is done by comparing the performance of the \Ac{BHH} to that of all the individual lower-level heuristics on all the datasets. These datasets include classification and regression problems of various sizes and complexity.  Detailed results are presented on the performance of the \Ac{BHH} as a selection mechanism for \acp{HH} and a brief comparison is made to other selection mechanisms. Finally, results are presented on the effects of the hyper-parameters of the \Ac{BHH} on the training process. Discussions follow on how the \Ac{BHH} is shown to automate the selection of the best heuristic during the training of \acp{FFNN}, as applied to a range of problems, alleviating the burden on researchers to apply traditional trial and error approaches. The chapter is concluded with a brief argument in favor of the general applicability of the \Ac{BHH} to more problem domains.

@@

The following chapter provides the reader with the background information required on \acp{ANN}.

>>>>

\begin{quotation}
      \noindent ``It is better to solve one problem five different ways, than to solve five problems one way.''
\end{quotation}
\begin{flushright}
      George Polya
\end{flushright}

\Ac{ML} is one of the most popular fields of research in \ac{AI} studies today. In recent years, \ac{ML} research has seen some notable achievements in academia \cite{ref:lecun:2015, ref:glorot:2010, ref:goodfellow:2014, ref:quoc:2017}, as well as the industry at large \cite{ref:silver:2016, ref:silver:2017, ref:zoph:2017, ref:lewis:2017}.  \ac{ML} research has grown tremendously over the past decade with successes like AlphaGo, which set new standards for \ac{AI} capabilities by beating the world's best Go player, Lee Sedol, 4-1 \cite{ref:san-hun:2016}.

Over the past few years, modern hardware capabilities have improved to the point where workloads in the field of \ac{ML} that were previously computationally infeasible, are now possible. One such sub-field of \ac{ML} is \acp{ANN}. \acp{ANN} can generally be described as well-organised structures of mathematical computation and are inspired by the biological brain \cite{ref:engelbrecht:2007}. \acp{ANN} can be trained, which is the equivalent of ``learning'' from data. Learning gives rise to the ability to apply some form of decision making. Decision making as an ability provides a wide-range of applications from healthcare to finance and yields great interest in the field. With the improvement of hardware came an influx of \ac{ML} researchers that focused their attention on training of \acp{ANN}.

\Acfp{FFNN} are specific types of \acp{ANN} \cite{ref:reed:1999}. A popular field of focus for studying \acp{ANN} is the process by which these models are trained.  The most common way of training \acp{FFNN} is \index{supervised learning}\textit{supervised learning}, which is a training technique that involves exposing the \acp{ANN} with input and comparing the produced output to that of predefined target data. Training of \acp{ANN} is seen as an optimisation problem. The \ac{ANN} maintains a set of parameters, referred to as ``weights'' and ``biases''. A search algorithm known as a \index{heuristic}\textit{heuristic} \cite{ref:pearl:1984} is used to assign optimal values to the parameters of the \ac{ANN}, such that a specified objective function is minimised. The research presented in this disseration focuses on the development of a specific type of heuristic, called a \acl{HH} to train \acp{FFNN} in a supervised learning approach.

This chapter provides the reader with a brief overview of the problem domain and outlines the research problem, objectives and purpose. The remainder of the chapter is presented as follows.

\begin{itemize}
      \item \textbf{Section \ref{sec:introduction:summary_research_domain}} provides the reader with brief summary of the research domain.

      \item \textbf{Section \ref{sec:introduction:problem}} outlines the research problem being addressed.

      \item \textbf{Section \ref{sec:introduction:motivation}} provides a motivation for this research.

      \item \textbf{Section \ref{sec:introduction:objectives}} presents the reader with the research objectives and outlines the goals of this dissertation.

@@

      \item \textbf{Section \ref{sec:introduction:outline}} provides a full dissertation outline.
\end{itemize}

@@

This section provides the reader with a brief summary of the research field and outlines key concepts that contribute to the topics discussed in this disseration.

@@

@@

Techniques exist to dynamically balance the trade-off between exploration and exploitation during the search process. An example of such a technique is to dynamically adjust and learn the heuristic \index{hyper-parameters}\textit{hyper-parameters} as part of the learning process. This field of study is known as \index{meta-learning}meta-learning \cite{ref:giraud:2004}. \index{meta-learning}Meta-learning of heuristic hyper-parameters as applied in the training of \acp{ANN} has shown to yield good generalisation results \cite{ref:hospedales:2020, ref:vilalta:2002}.

The dynamic nature of the learning process leads towards the idea that \ac{ML} models could be trained in a way such that the learning process and learning mechanism applied are not statically defined, but rather dynamic and under the control of some mechanism.

A recent suggestion related to the field of \index{meta-learning}meta-learning is to dynamically select and/or adjust the heuristic used throughout the training process. This approach focuses on the hybridisation of learning paradigms. The main concept behind this paradigm is that the learning process is dynamic and problem specific. A particular learning technique might work well for one problem and not for another. At the same time, a particular learning technique might work well for a particular part of the search landscape, but not for another. By dynamically combining the best of different learning paradigms throughout the learning process, a trade-off can be made between exploration and exploitation as is required.

One such form of hybridisation of learning paradigms is referred to as \textit{heterogeneous learning approaches}. Heterogeneous learning approaches make use of different \textit{search behaviours} by selecting from a behaviour pool. Heterogeneous approaches have shown to balance the trade-off between exploration and exploitation \cite{ref:nepomuceno:2013}.

A step further in the concept of hybridisation of learning paradigms is that of hybridisation of different \textit{heuristics} as they are applied to some optimisation problem \cite{ref:burke:2013}. These methods are referred to as \acfp{HH} and focus on finding the best heuristic in \textit{heuristic-space} to solve a specific problem. One such form of \ac{HH} is a population-based approach that guides the search process by automatically selecting heuristics from a heuristic-pool to be applied to a collection of different candidate solutions in the solution-space. This collection of candidate solutions are referred to as a \textit{population} of \textit{entities}, where each \textit{entity} is a single candidate solution to the problem being optimised. Population-based \acp{HH} implement a strategy where multiple heuristics work together to solve a problem. This technique requires a mechanism that selects a specific heuristic to be applied to a specific candidate solution.

Finding the best heuristic to use is non-trivial and some \textit{selection strategy} must be used to select the best heuristic. \acp{HH} that implement such a selection strategy are referred to as \textit{selection} \acp{HH}. The term \textit{selection} refers to the ability of the \ac{HH} to select the best heuristic from a pool of low-level heuristics.

One specific type of selection \ac{HH} is called a \index{multi-method populated-based meta-heuristic}\textit{multi-method population-based \index{meta-heuristic}meta-heuristic} \cite{ref:vanderstockt:2018}. The term \index{multi-method}\textit{multi-method} refers to the incorporation of different low-level heuristics, with different search behaviours, into the heuristic-space. The term \textit{population-based} refers to the utilisation of a population of entities that represent candidate solutions. Finally, the term \index{meta-heuristic}\textit{meta-heuristics} refers to the \ac{HH} as a heuristic that does not have any domain knowledge and only makes use of information from the search process.

\citeauthor{ref:grobler:2015} \cite{ref:grobler:2015} mentioned that \acp{HH} have been shown to solve a number of problems including bin-packing, examination timetabling, production scheduling, the traveling salesman problem, vehicle routing problem and many more.

The following section outlines the research problem that this dissertation aims to address and is followed by a motivation for this research.

@@

Many different heuristics have been developed and used to train \acp{FFNN} \cite{ref:gudise:2003, ref:rakitianskaia:2012, ref:montana:1989}. Each of these heuristics have different search behaviours, characteristics, strengths and weaknesses. Finding the best heuristic to train the \ac{FFNN} is required in order to yield optimal results. This process is often non-trivial and could be a time-consuming exercise.  Consider that selection of the best heuristic as applied to optimisation problems, such as training \acp{FFNN}, is problem dependent \cite{ref:allen:1996, ref:drake:2020, ref:pillay:2018}.

Careful, systematic selection is thus required to find and select the best heuristic to train \acp{FFNN}. In the past, researchers selected the best heuristic by trail and error. A set of heuristics and carefully selected hyper-parameters would be implemented, followed by an empirical test to evaluate the performance of each heuristic for a given problem domain \cite{ref:pillay:2015}. In this way, researchers were able to determine which heuristics and which hyper-parameters performed well for different problems. However, this approach is problematic, because it is time-consuming and laborious.

@@

@@

In Section \ref{sec:introduction:problem} above, it is mentioned that finding the best heuristic to train \acp{FFNN} is a timely and tedious process. A modern approach is to use \acp{HH} to automate the process of selecting the best heuristic when applied to some optimisation problem. The best heuristic might not be a single heuristic, but rather a hybridisation of heuristics \cite{ref:pillay:2015}.

In the general context of optimisation, many different types of \acp{HH} have been implemented and applied to many different problems. Some notable examples include the \index{simulated annealing}simulated annealing-based \ac{HH} by \citeauthor{ref:dowsland:2007} \cite{ref:dowsland:2007}, the \index{tabu-search}tabu-search \ac{HH} of \citeauthor{ref:burke:2010} \cite{ref:burke:2010}, the \index{heterogeneous meta-hyper-heuristic}heterogeneous meta-hyper-heuristic by \citeauthor{ref:grobler:2012} \cite{ref:grobler:2012} and work done by \citeauthor{ref:vanderstockt:2018} \cite{ref:vanderstockt:2018} on the analysis of selection hyper-heuristics for population-based meta-heuristics in real-valued dynamic optimization. However, research on the application of \acp{HH} in the context of \acs{FFNN} training is scarce. \citeauthor{ref:nel:2021} \cite{ref:nel:2021} provides the first research in this field, applying a \acs{HH} to \acs{FFNN} training.

\acp{FFNN} training is seen as a computational search problem. \Acl{HH} as defined by Burke et al. \cite{ref:burke:2010} is ``a search method or learning mechanism for selecting or generating heuristics to solve computational search problems''. \acp{HH} is a field of research aimed at the automated selection, combination, adaptation or generation of multiple lower-level heuristics to efficiently solve computational search problems.

@@

This research takes a particular interest in developing a selection \acs{HH} that makes use of probability theory and Bayesian statistical concepts to guide the heuristic selection process. This research develops the novel \Acf{BHH}, a new high-level heuristic that utilises a statistical approach, referred to as \index{Bayesian analysis}\textit{Bayesian analysis}, which combines prior information with new evidence to the parameters of a selection probability distribution. This selection probability distribution is the mechanism by which the \ac{HH} selects appropriate heuristics to train \acp{FFNN} during the training process. This process takes place dynamically and during the training process (online learning).

@@

The main objective of this research is focused on developing a novel \Ac{BHH} selection mechanism in a \ac{HH} framework that can be used to train \acp{FFNN}. In order to reach this objective, the following sub-objectives are defined.

@@

      \item Conduct a literature study on different types of heuristics that have been used to train \acp{FFNN}. The literature study will provide the necessary background information to understand how different heuristics can be used in the set of heuristics to be selected to train the \ac{FFNN}.

@@

      \item Conduct an empirical study to investigate the behavioural characteristics of the \ac{BHH} as it is used to train \acp{FFNN} on an example problem.

      \item Conduct an empirical study and critically evaluate the performance of the developed \Ac{BHH} compared to individual heuristics in the heuristic-space as they are used to train \acp{FFNN} on a number of different problems.

@@

By achieving the objectives outlined above, a number of research contributions are made. These contributions are presented in this following section.

@@

The results obtained from this research contribute to the field of study in the
following ways.

@@

      \item The results of the empirical study show that the \Ac{BHH} is able to select the best heuristic to train \ac{FFNN} in general. This relieves researchers from the burden of having to do this selection process manually through trial and error.

@@

The remainder of this dissertation is structured as follows.

\begin{itemize}
      \item \textbf{Chapter~\ref{chap:anns}} provides a literature study on \acp{ANN} and the various components that make up an \acl{AN}. The focus is on training of \acp{FFNN}. It is shown how training of \acp{FFNN} is seen as an optimisation problem.

@@

      \item \textbf{Chapter~\ref{chap:methodology}} presents a detailed description of the empirical process and the setup of each experiment. It discusses the datasets used, the \ac{FFNN} architecture and topology used, heuristics that are used, the configuration of hyper-parameters, initialisation techniques used and performance measures used.  Finally, discussions follow on how the results are analysed. This includes a discussion on the method used to determine statistical significance.

      \item \textbf{Chapter~\ref{chap:results}} provides and discusses the results of the empirical study in detail. A baseline comparison is done by comparing the performance of the \Ac{BHH} to that of all the individual lower-level heuristics on all datasets. These datasets include classification and regression problems of various sizes and complexity.  Detailed results are presented on the performance of the \Ac{BHH} as a selection mechanism for \acp{HH} and a brief comparison is made to other selection mechanisms. Finally, results are presented on the effects of the hyper-parameters of the \Ac{BHH} on the training process. Discussions follow on how the \Ac{BHH} is shown to automate the selection of the best heuristic during the training of \acp{FFNN}, as applied to a range of problems, alleviating the burden on researchers to apply traditional trial and error approaches. The chapter is concluded with a brief argument in favor of the general applicability of the \Ac{BHH} to more problem domains.

@@

### 2022-10-09 dissertation/chapters/03-heuristics.tex
✨ (dissertation): Chapter 1 done

\begin{equation}
    \label{eq:heuristics:pso:velocity_clamping}
    \begin{split}
        v_{ij}(t+1)= 
        \begin{cases}
            v'_{ij}(t+1) & \text{if } |v'_{ij}(t+1)| < V_{max,j} \\
            -V_{maxj}   & \text{if } v'_{ij}(t+1) \leq -V_{max,j}\\
            V_{maxj}   & \text{if } v'_{ij}(t+1) \geq V_{max,j}
        \end{cases}
    \end{split}
\end{equation}

@@

This section aims to introduce the reader to the next population-based \ac{MH} called \acl{DE}. Similar to \ac{PSO}, \Ac{DE} is a stochastic population-based search strategy developed by Storm and Price \cite{ref:price:2006} in 1995. \Ac{DE} shares a lot of similarities with other evolutionary \ac{MH} paradigms such as \acp{PSO} and \acp{GA}. However, \ac{DE} differs significantly in the sense that it makes use of distance and direction information from the current population to guide the search process \cite{ref:engelbrecht:2007}. Originally, \ac{DE} was focussed on multi-dimensional real-values optimisation problems, but unlike gradient-based heuristics, it does not require any gradient information. This means that \ac{DE} does not require the underlying optimisation problem to be differentiable, meaning that it can be applied to problems that are discrete, noisy and dynamic \cite{ref:rocca:2011}.

@@

\citeauthor{ref:engelbrecht:2007}\cite{ref:engelbrecht:2007} mentions that crossover operators can be further classified based on their encoding/representation scheme. These include binary-specific operators used for binary representations and operators focussed on floating-point representations. Since the focus is put on training \acp{FFNN}, from this point on, floating-point representations are assumed.

>>>>

\begin{equation}
      \label{eq:heuristics:pso:velocity_clamping}
      \begin{split}
            v_{ij}(t+1)=
            \begin{cases}
                  v'_{ij}(t+1) & \text{if } |v'_{ij}(t+1)| < V_{max,j}   \\
                  -V_{maxj}    & \text{if } v'_{ij}(t+1) \leq -V_{max,j} \\
                  V_{maxj}     & \text{if } v'_{ij}(t+1) \geq V_{max,j}
            \end{cases}
      \end{split}
\end{equation}

@@

This section aims to introduce the reader to the next population-based \ac{MH} called \acl{DE}. Similar to \ac{PSO}, \Ac{DE} is a stochastic population-based search strategy developed by Storm and Price \cite{ref:price:2006} in 1995. \Ac{DE} shares a lot of similarities with other evolutionary \ac{MH} paradigms such as \acp{PSO} and \acp{GA}. However, \ac{DE} differs significantly in the sense that it makes use of distance and direction information from the current population to guide the search process \cite{ref:engelbrecht:2007}. Originally, \ac{DE} was focused on multi-dimensional real-values optimisation problems, but unlike gradient-based heuristics, it does not require any gradient information. This means that \ac{DE} does not require the underlying optimisation problem to be differentiable, meaning that it can be applied to problems that are discrete, noisy and dynamic \cite{ref:rocca:2011}.

@@

\citeauthor{ref:engelbrecht:2007}\cite{ref:engelbrecht:2007} mentions that crossover operators can be further classified based on their encoding/representation scheme. These include binary-specific operators used for binary representations and operators focused on floating-point representations. Since the focus is put on training \acp{FFNN}, from this point on, floating-point representations are assumed.
